---
title: "STAT 478 Project"
author: "Logan Rosentreter & Zachary Kern"
date: "Mar 19, 2023"
header-includes:
  - \usepackage{pdfpages}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r package}
# check if `tidyr` package is installed;
# otherwise, install and load if 'tidyr' is absent
if(!require('tidyr'))
install.packages('tidyr', repos = 'https://cloud.r-project.org')

# load libraries

#rename
library(dplyr)
```

---

# Overview

The 10-Year Real Interest Rate is a commonly used economic indicator that reflects the average annual return on investments adjusted for inflation over a 10-year period. The Fred STL dataset tracks the 10-Year Real Interest Rate in the United States, providing valuable insights into the state of the economy and the financial market. The 10-Year Real Interest Rate is an important indicator for investors, policymakers, and economists because it provides a measure of the real cost of borrowing and the expected return on investment. When the 10-Year Real Interest Rate is low, it can stimulate economic growth by making borrowing cheaper and encouraging investment. Conversely, when the 10-Year Real Interest Rate is high, it can restrict economic growth by increasing the cost of borrowing and reducing investment.

```{r}
# read in data set
int_rate <- read.csv("RIRA10Y.csv")

# rename columns for easier data manipulation
int_rate <- int_rate %>% 
    rename("INT_RATE_10Y" = REAINTRATREARAT10Y)

# convert date to type(date)
int_rate <- int_rate %>%
    mutate(DATE = as.Date(DATE, sep="-", "%Y-%m-%d"))
str(int_rate)

int.ts <- ts(int_rate$INT_RATE_10Y)
plot(int.ts)
int_acf=acf(int_rate, lag.max=25,type="correlation",main="ACF of Project Data")
```

Testing first change
1. Problem definition

2. Data description

3. Data Analysis

4. Model specification and fitting

5. Model validation and diagnostics

6. Forecasting
