---
title: "STAT 478 Project"
author: "Logan Rosentreter & Zachary Kern"
date: "Mar 19, 2023"
bibliography: source.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

```{r funtions, include=F}
# Define a function variogram
variogram <- function (x, lag) {
  Lag <- NULL
  var_k <- NULL
  vario <- NULL
  for (k in 1:lag){
    Lag[k] <- k
    var_k[k] = (sd(diff(x,k)))^2
    vario[k] = var_k[k] / var_k[1]
  }
  return (as.data.frame( cbind (Lag , vario )))
}
```

```{r package, include=F}
# check if `tidyr` package is installed;
# otherwise, install and load if 'tidyr' is absent
if(!require('prophet'))
install.packages('prophet', repos = 'https://cloud.r-project.org')

# load libraries

#rename
library(dplyr)

# acf2
library(astsa)

# auto.arima
library(forecast)

# prophet
library(prophet)
```

---

# Things to include

- plot, acf, variogram, (decompose)
- qqplot, aug dick full,
- white noise test : box-pierce
- normality test : anderson-darling
- auotcorrelation & time series regression : dwt
- auotcorrelation & regression : cochrane orcutt
- auotcorrelation : Ljung-Box

## general approach

1. plot 

  - determine basic features

    - trend, season, outliers

2. elimin trend, seas

  - diff

  - apply appropriate model

3. develop forecast model for residuals

4. validate performance

- split-sample

- cross-validation
 
5. find diff b/n orig and forecast / smoothed

6. find prediction intervals of forecast

7. develop procedure for detecting deterioration in forecast, quickly

# also want to use prophet 

## evaluation

- ME, MAD, MSE, MPE, MAPE

# 1. Problem definition / Introduction

The Fred STL dataset tracks the [10-Year Real Interest Rate](https://fred.stlouisfed.org/series/REAINTRATREARAT10Y) in the United States [@fred_10yr_intrate]. The 10-Year Real Interest Rate provides valuable insights into the state of the economy and the financial market, as it provides a measure of the real cost of borrowing and the expected return on investment. When the 10-Year Real Interest Rate is low, it can stimulate economic growth by making borrowing cheaper and encouraging investment. When the 10-Year Real Interest Rate is high, it can restrict economic growth by increasing the cost of borrowing and reducing investment.

```{r}
# read in data set
int.rate <- read.csv("RIRA10Y.csv")

# rename columns for easier data manipulation
int.rate <- int.rate %>% 
    rename("INT_RATE_10Y" = REAINTRATREARAT10Y)

# convert date to type(date)
int.rate <- int.rate %>%
    mutate(DATE = as.Date(DATE, sep="-", "%Y-%m-%d"))
str(int.rate)

int.ts <- ts(int.rate$INT_RATE_10Y, freq=12)
plot(decompose(int.ts))
plot(int.ts)
int_acf2 <- acf2(int.ts, max.lag = 50)
plot(variogram(int.ts, 50), pch=19, cex=0.65)
```

## notes - orig

* data shows down trend, no seas
* noncons mean, cons variance
* acf -> nonstation
* variogram -> monotonically increasing for long period of time -> nonstat

```{r}
int.lm <- lm(int.ts ~ int.rate$DATE)
summary(int.lm)
anova(int.lm)
```


```{r}
# differenced data
dint.ts <- diff(int.ts)
plot(dint.ts)
dint.acf2 <- acf2(dint.ts)
plot(variogram(dint.ts, 50), pch=19, cex=0.65)
```

## notes - diff

* data shows no trend, no seas, rand scatter
* cons mean, cons variance
* acf -> station
* variogram -> random scatter -> stat

```{r}
dint.lm <- lm(dint.ts ~ int.rate$DATE[-1])
summary(dint.lm)
anova(dint.lm)
```

```{r holtwinters}
int.ses <- HoltWinters(int.ts)
plot(int.ses)
int.ses.for <- forecast(int.ses, h=12)
plot(int.ses.for)
```


```{r arima}
int.arima <- auto.arima(int.ts)
plot(forecast(int.arima))
```

```{r prophet}
int.rate.prophet <- int.rate %>%
    rename(ds = DATE, y = INT_RATE_10Y)

# implementing prophet
m <- prophet(int.rate.prophet)
future <- make_future_dataframe(m, freq=12, periods = 12)
forecast <- predict(m, future)
plot(m, forecast)
prophet_plot_components(m, forecast)
```



## notes - log trans

* cant use logarithm, because of negatives

# 2. Data description

# 3. Data Analysis

# 4. Model specification and fitting

# 5. Model validation and diagnostics

# 6. Forecasting

# Conclusion

--

